RegressionPvNP

### 2017-11-03 RegressionPvNP.Rmd
updating glm pipeline

@@

@@

```{r labelsPNP, echo=F, message=F, include=T}
## binomial: dysplasia 1, BE 0
labels = unlist(sapply(patient.data, function(df) {
  df$info = arrange(df$info, Endoscopy.Year, Pathology)
  label = as.integer(df$info$Status == 'P') #as.integer(df$info$Pathology %in% c('HGD', 'IMC'))
  names(label) = df$info$Samplename
  return(label)
}))
names(labels) = sub('.*\\.', '',  names(labels))

@@

dysplasia.df = t(mergedDf[,names(labels)])
dim(dysplasia.df)

@@

```{r cv.funcs, message=F, warning=F, echo=T, fig.height=6, fig.width=6}
pi.hat<-function(x) exp(x)/(1+exp(x))

non.zero.coef<-function(fit, s) {
  cf =  as.matrix(coef(fit, s))
  cf[which(cf != 0),][-1]
}

coef.stability<-function(opt, nz.list) {
  for (i in 1:length(nz.list)) {
    df = as.data.frame(nz.list[[i]])
    colnames(df) = i
    opt = merge(opt, df, by='row.names', all.x=T)
    rownames(opt) = opt$Row.names
    opt$Row.names = NULL
    opt[,(i+1)] = as.integer(!is.na(opt[,(i+1)]))
  }
  opt = opt[order(sapply(rownames(opt), function(x) as.numeric(unlist(strsplit(x, ':'))[1])  )),]
  return(opt)
}

create.patient.sets<-function(pts, n, splits, minR=0.2) {
  # This function just makes sure the sets don't become too unbalanced with regards to the labels.
  check.sets<-function(df, grpCol, min) {
    sets = table(cbind.data.frame('set'=df[[grpCol]], 'labels'=labels[df$Samplename]))
    while ( (length(which(sets/rowSums(sets) < minR) ) >= 2 | length(which(sets/rowSums(sets) == 0)) > 0) ) {
      #print(sets/rowSums(sets))
      s = sample(rep(seq(5), length = length(unique(df$Hospital.Research.ID))))
      df2 = merge(df, cbind('Hospital.Research.ID'=unique(df$Hospital.Research.ID), 'tmpgrp'=s), by="Hospital.Research.ID")
      df[[grpCol]] = df2$tmpgrp
      sets = table(cbind.data.frame('set'=df[[grpCol]], 'labels'=labels[df$Samplename]))
    }
    return(df[[grpCol]])
  }

  s = sample(rep(seq(splits), length = length(unique(pts$Hospital.Research.ID))))
  patients = merge(pts, cbind('Hospital.Research.ID'=unique(pts$Hospital.Research.ID), 'group'=s), by="Hospital.Research.ID")  
  colnames(patients)[3] = c('fold.1')
  patients$fold.1 = check.sets(patients, 'fold.1', minR)

  for (i in 2:n) {
    s = sample(rep(seq(5), length = length(unique(pts$Hospital.Research.ID))))
    patients = merge(patients, cbind('Hospital.Research.ID'=unique(patients$Hospital.Research.ID), 'group'=s), by="Hospital.Research.ID")  
    foldcol = grep('group',colnames(patients))
    colnames(patients)[foldcol] = paste('fold',i,sep='.')
    patients[[paste('fold',i,sep='.')]] = check.sets(patients, paste('fold',i,sep='.'))
  }

  return(patients)
}

binomial.deviance<-function(pmat, y) {
  # Binomial deviance, lifted from cv.lognet
  prob_min = 1e-05; prob_max = 1 - prob_min
  pmat = pmin(pmax(pmat, prob_min), prob_max)
  dev = apply(pmat, 2, function(p)  -2*((y==1)*log(p)+(y==0)*log(1-p)) )
  return(dev)
}

crossvalidate.by.patient<-function(x,y,lambda,pts,a=1,nfolds=10, splits=5, fit=NULL, minR=0.2, select='deviance', ...) {
  if (nfolds > 5) minR = 0.1
  x = as.matrix(x)
  message(paste("Running", splits, "splits",nfolds,"times on", paste(dim(x), collapse=':'), 'alpha=',a ))
  fit.e = list()

  if (ncol(pts) == (nfolds+2)) {
    tpts = pts
  } else {
    tpts = create.patient.sets(pts, nfolds, splits, minR)
  }

  cv.pred = (matrix(nrow=0, ncol=length(lambda)))
  cv.binomial.deviance = (matrix(nrow=0, ncol=length(lambda)))
  for (n in 1:nfolds) {  
    message(paste(n, "fold"))
    setCol = grep(paste('^fold.',n,'$',sep=''), colnames(tpts))

    cv.class = matrix(nrow=splits, ncol=length(lambda))
    deviance = matrix(nrow=splits, ncol=length(lambda))
    for (i in 1:splits) { 
      message(paste(i, "split"))
      test.rows = which(rownames(x) %in% tpts[which(tpts[,setCol] == i), 'Samplename'])
      test = x[test.rows,]
      training = x[-test.rows,]
      # pre-spec lambda seq
      fitCV = glmnet(training, y[-test.rows], lambda=lambda, alpha=a, family='binomial', standardize = F) 
      # autoplot(fit) + theme(legend.position="none")

      # Confusion matrix: quantitiative, pos results + neg results / number of test rows
      pred <- pi.hat(predict(fitCV, test, type='link'))  # pi.hat turns these into probabilities from logit
      cv.class[i,] = apply(pred, 2, function(p) {
        (p%*%y[test.rows] + (1-p) %*% (1-y[test.rows]))/length(test.rows)
      })

      # Binomial deviance, lifted from cv.lognet
      dev = binomial.deviance(predict(fitCV, test, type='response'), as.factor(y[test.rows]))
      deviance[i,] = apply(dev, 2, weighted.mean, w=rep(1, nrow(dev)), na.rm=T)

      fit.e[[length(fit.e)+1]] = fitCV
    }
    cv.pred = rbind(cv.pred, cv.class)    
    cv.binomial.deviance = rbind(cv.binomial.deviance, deviance)
  }

  # Not really sure what to do with the binomial deviance now...
  df = cbind.data.frame('lambda.at'=1:ncol(cv.pred),
                        'mean'= colMeans(cv.pred), 
                        'sme'= apply(cv.pred, 2, sd)/sqrt(nrow(cv.pred)), 
                        'sd'= apply(cv.pred, 2, sd), 
                        'lambda'= lambda,
                        'log.lambda' = log(lambda),
                        'mean.b.dev' = colMeans(cv.binomial.deviance),
                        'sd.b.dev' = apply(cv.binomial.deviance, 2,sd),
                        'sme.b.dev' = apply(cv.binomial.deviance, 2, sd)/sqrt(nrow(cv.binomial.deviance)),
                        'lambda.min' = F, 'lambda.1se' = F)

  if (select == 'classification') {
    # Minimize the classification mean error & select min lambda
    df[which.max(subset(df, sme < median(sme) & sme > min(sme))$mean), 'lambda.min'] = T

    #df$lambda.min = df$mean == max(df$mean)
    se1 = df[df$lambda.min == T, 'mean'] - df[df$lambda.min == T, 'sme']

    lambda.1se.search = subset(df, mean < se1 & !lambda.min & log.lambda > subset(df, lambda.min)$log.lambda) 
    if (nrow(lambda.1se.search) > 0) {
      arrange(lambda.1se.search, -mean)[1:10,]
      df[arrange(subset(df, mean < se1 & !lambda.min & log.lambda > subset(df, lambda.min)$log.lambda ), -mean)[1, 'lambda.at'], 'lambda.1se'] = T
    }
    #df[df$log.lambda < subset(df, mean == max(mean))$log.lambda-sd(df$log.lambda),][1, 'lambda-1se'] = TRUE
  } else if (select == 'deviance') {
    df[which.min(df$mean.b.dev), 'lambda.min'] = T

    lmin = subset(df, lambda.min)$lambda
    lambdas = subset(df, log.lambda <= log(lmin) + log(sd(lambda))/2 )
    n = 3
    while (nrow(lambdas) <= 0 & n < 10) {
      lambdas = subset(df, log.lambda <= log(lmin) + log(sd(lambda))/n )
      n = n+1
    }
    if (nrow(lambdas) <= 0) {
      df[subset(df, log.lambda <= log(lmin) )[1, 'lambda.at'], 'lambda.1se'] = T  
    } else {
      df[subset(df, log.lambda <= log(lmin) + log(sd(lambda))/n )[1, 'lambda.at'], 'lambda.1se'] = T  
    }
  }

  lambda.min = lambda[subset(df, lambda.min == T)$lambda.at] 
  lambda.1se = lambda[subset(df, lambda.1se == T)$lambda.at]

  nzcf = lapply(fit.e, non.zero.coef, s=lambda.1se)

  plots = plot.patient.cv(df, fit)
  plots$performance = plots$performance + theme(legend.position='bottom') + scale_colour_discrete(name = "") +
    labs(title=paste(splits,' splits, ',nfolds,' folds, alpha=',a, sep=''), y='mean pred.', x='log(lambda)') 
  plots$deviance = plots$deviance + theme(legend.position='bottom') + scale_colour_discrete(name="") + 
    labs(title=paste(splits,' splits, ',nfolds,' folds, alpha=',a, sep=''), y='mean Binomial Deviance', x='log(lambda)')

  return(list('max.cm'=df[df$lambda.1se ==T,'mean'], 
              'lambda.min'=lambda.min, 'lambda.1se'=lambda.1se, 'lambdas'=df, 'plot'=plots$performance, 'deviance.plot'=plots$deviance, 'non.zero.cf'=nzcf))
}

plot.patient.cv<-function(df, fit=NULL) {
  gp = ggplot(df, aes(y=mean,x=log.lambda)) + geom_point() + geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + 
    geom_point(data=subset(df, lambda.min == T), aes(y=mean, x=log.lambda, colour="lambda.min"), size=2 ) +  
    geom_vline(xintercept = subset(df, lambda.min == T)$log.lambda, colour='grey') +
    annotate("text", x=subset(df, lambda.min == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
             label=paste(round(df[subset(df, lambda.min == T)$lambda.at,c('mean', 'sme')],3), collapse='\n+/-'))

  gpD = ggplot(df, aes(y=mean.b.dev,x=log.lambda)) + geom_point() +
    geom_errorbar(aes(ymin=mean.b.dev-sme.b.dev, ymax=mean.b.dev+sme.b.dev)) + 
    geom_point(data=subset(df, lambda.min == T), aes(y=mean.b.dev, x=log.lambda, colour="lambda.min"), size=2 ) +          
    geom_vline(xintercept = subset(df, lambda.min == T)$log.lambda, colour='grey') +
    annotate("text", x=subset(df, lambda.min == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
             label=paste(round(df[subset(df, lambda.min == T)$lambda.at,c('mean', 'sme')],3), collapse='\n+/-'))

  if (nrow(subset(df, lambda.1se == T)) > 0) {
    gp = gp + geom_point(data=subset(df, lambda.1se == T), aes(y=mean, x=log.lambda, colour="lambda.1se"), size=2 ) +
      geom_vline(xintercept = subset(df, lambda.1se == T)$log.lambda, colour='grey') +
      annotate("text", x=subset(df, lambda.1se == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
               label=paste(round(df[subset(df, lambda.1se == T)$lambda.at,c('mean', 'sme')],3), collapse='\n+/-'))

    gpD = gpD + geom_point(data=subset(df, lambda.1se == T), aes(y=mean.b.dev, x=log.lambda, colour="lambda.1se"), size=2 ) +
      geom_vline(xintercept = subset(df, lambda.1se == T)$log.lambda, colour='grey') +
      annotate("text", x=subset(df, lambda.1se == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
               label=paste(round(df[subset(df, lambda.1se == T)$lambda.at,c('mean', 'sme')],3), collapse='\n+/-')) 
  }

  gp = gp + theme(legend.position='bottom') + scale_colour_discrete(name = "") 
  gpD = gpD + theme(legend.position='bottom') + scale_colour_discrete(name="")

  if (!is.null(fit)) {
    df$nzcoef = sapply(df$lambda, function(l) length(non.zero.coef(fit, l)))

    coef.min = subset(df, lambda.min == T)$nzcoef
    coef.1se = subset(df, lambda.1se == T)$nzcoef

    d = coef.1se-coef.min

    coef.text = arrange(subset(df,nzcoef %in% c(0,  coef.min, coef.1se, max(nzcoef) )), -lambda.min, -lambda.1se)
    coef.text = arrange(coef.text[!duplicated(coef.text$nzcoef),], -nzcoef)

    gp = gp + annotate("text", y=min(df$mean), x=coef.text$log.lambda, label=coef.text$nzcoef, color='darkblue')
    gpD = gpD + annotate("text", y=max(df$mean.b.dev), x=coef.text$log.lambda, label=coef.text$nzcoef, color='darkblue')
  }
  return(list('performance'=gp, 'deviance'=gpD))
}

@@

coefs = list(); plots = list(); performance.at.1se = list(); models = list();
folds = 10; splits = 5 
alpha.values = c(0, 0.5,0.7,0.8,0.9,1)

file = paste(data, 'Analysis/patient_folds.tsv', sep='/')
if (file.exists(file)) {
  message(paste("Reading folds file", file))
  sets = read.table(file, header=T, sep='\t')
} else {
  #sets = create.patient.sets(pts, folds, splits, 0.2)  
  #write.table(sets, quote=F, sep='\t', row.names=F, file=file)
  stop("Missing patient folds file")
}

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=T)
} else {
  for (a in alpha.values) {
    fit0 <- glmnet(dysplasia.df, labels, alpha=a, nlambda=nl, family='binomial', standardize=F) # all patients
    #autoplot(fit0, xvar='lambda', main=paste('fit0, all samples, alpha=',a,sep='')) + theme(legend.position='none')

    l = fit0$lambda
    if (a == 0) {
      l = sort(c(fit0$lambda, seq(exp(-5), exp(-10), -1e-6),
            seq(exp(-10), exp(-15), -1e-8),
            seq(exp(-15), exp(-20), -1e-9)), decreasing=T)
    }
    cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=l, pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', standardize=F)
    #cv.patient$deviance.plot

    lambda.opt = ifelse( length(cv.patient$lambda.1se) > 0, cv.patient$lambda.1se, cv.patient$lambda.min )

@@

    performance.at.1se[[as.character(a)]] = subset(cv.patient$lambdas, lambda.1se == T)
    models[[as.character(a)]] = fit0
  }
  save(plots, coefs, performance.at.1se, dysplasia.df, models, labels, file=file)
}

@@

if (file.exists(file)) {
  load(file, verbose=T)
} else {
  performance.at.1se = c(); coefs = list(); plots = list(); fits = list()
  # Remove each patient (LOO)
  for (pt in names(pg.samp)) {
    print(pt)
    tmp.patient.data = patient.data[subset(sum.patient.data, Hospital.Research.ID != pt)$Hospital.Research.ID]
    samples = as.vector(unlist(sapply(tmp.patient.data, function(df) df$info$Samplename )))

@@

    # Fit generated on all samples, including HGD
    a = select.alpha
    fitLOO <- glmnet(training, labels[train.rows], alpha=a, family='binomial', nlambda=nl, standardize=F) # all patients
    #l = sort(c(fitLOO$lambda, seq(exp(-5), exp(-10), -0.001),
    #      seq(exp(-10), exp(-15), -1e-7), 
    #      seq(exp(-15), exp(-20), -1e-9)), decreasing=T)
    l = fitLOO$lambda
    cv = crossvalidate.by.patient(x=training, y=labels[train.rows], lambda=l, 
                                  pts=subset(pts, Samplename %in% samples), a=a, nfolds=folds, splits=splits, fit=fitLOO, standardize=F)
    #plots[[pt]] = cv$plot + ggtitle(pt)

@@

ggplot(as.data.frame(performance.at.1se), aes(y=performance.at.1se, x='LOO')) + ylim(0.5,0.9) +
  geom_boxplot( fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA) + geom_jitter() +
  labs(y='performance at lambda.1se', x='', title='Performance for Leave One (patient) Out, at alpha=1')

@@

```{r, warning=F, message=F, fig.height=4, fig.width=4, eval=F}
### How many samples?  TBD

#Clinicians asked how many samples they needed to take.

# How do I evaluate the predictions?

#Basically I need to to a FNR and FPR calculation that is extremely conservative and use that somehow

all = do.call(rbind,lapply(pg.samp, function(df) {
  df[,c('Status','Prediction','Prediction.Dev.Resid')]
}))

ggplot(all, aes(Prediction.Dev.Resid, Prediction, color=Status)) + geom_point()  + 
  labs(y='Predicted (response)', x='Binomial dev residual?')
```

@@

```{r, eval=F}
a=1

plot.cv<-function(cv, title='') {
  grid.arrange( arrangeGrob(cv$plot+ggtitle('Classification'), cv$deviance.plot+ggtitle('Binomial Deviance'), top=title, ncol=2) )
}

more.l<-function(lambda) {
    l = sort(c(lambda, seq(exp(-5), exp(-10), -1e-6),
            seq(exp(-10), exp(-15), -1e-8),
            seq(exp(-15), exp(-20), -1e-9)), decreasing=T)
}

# random
df11 = dysplasia.df[,grep('^11:', colnames(dysplasia.df))]
fit0 <- glmnet(df11, labels, alpha=a, nlambda=nl, family='binomial', standardize=F) # all patients
cv11 = crossvalidate.by.patient(x=df11, y=labels, lambda=more.l(fit0$lambda), pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', standardize=F)
plot.cv(cv11,"Chr11")

df17 = dysplasia.df[,grep('^17:', colnames(dysplasia.df))]
fit0 <- glmnet(df17, labels, alpha=a, nlambda=nl, family='binomial', standardize=F) # all patients
cv17 = crossvalidate.by.patient(x=df17, y=labels, lambda=more.l(fit0$lambda), pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', standardize=F)
plot.cv(cv17, 'Chr17')

## No HGD
df17 = dysplasia.df[subset(patient.info, Pathology %in% c('BE','ID','LGD'))$Samplename,grep('^17:', colnames(dysplasia.df))]
fit0 <- glmnet(df17, labels[subset(patient.info, Pathology %in% c('BE','ID','LGD'))$Samplename], alpha=a, nlambda=5000, family='binomial', standardize=F)
cv17 = crossvalidate.by.patient(x=df17, y=labels[subset(patient.info, Pathology %in% c('BE','ID','LGD'))$Samplename], lambda=fit0$lambda, pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', standardize=F)
plot.cv(cv17, 'Chr17, no HGD')

df3 = dysplasia.df[,grep('^3:', colnames(dysplasia.df))]
fit0 <- glmnet(df3, labels, alpha=a, nlambda=nl, family='binomial', standardize=F) # all patients
cv3 = crossvalidate.by.patient(x=df3, y=labels, lambda=fit0$lambda, pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', standardize=F)
plot.cv(cv3, 'Chr3')

>>>>

## update the info as it has changed
patient.data = lapply(patient.data, function(df) {
  df$info = subset(patient.info, Patient == unique(df$info$Patient))
  df
})

validation.patient.data = lapply(validation.patient.data, function(df) {
  df$info = subset(validation.patient.info, Patient == unique(df$info$Patient))
  df
})

@@

```{r, rawvalues, echo=F, eval=F}

p = ggplot(chr.lengths, aes(x=genome.length))

for (pt in subset(sum.patient.data, Status == 'P')$Hospital.Research.ID) {
  df = merge(patient.data[[pt]]$seg.vals, chr.lengths[,c('chrom','chr.length','genome.length')])
  df$g.start = df$start.pos; df$g.end = df$end.pos
  df[which(df$chrom != 1), 'g.start'] = rowSums(df[which(df$chrom != 1), c('start.pos', 'genome.length')])
  df[which(df$chrom != 1), 'g.end'] = rowSums(df[which(df$chrom != 1), c('end.pos', 'genome.length')])

  df = melt(df, measure.vars=grep('^D',colnames(df),value=T))

  p = p + geom_segment(data=df, aes(x=g.start,xend=g.end,y=value, yend=value,color=variable),show.legend=F)  
}

p1 = ggplot(chr.lengths, aes(x=genome.length)) + ylim(0,3)
for (pt in subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID) {
  df = merge(patient.data[[pt]]$seg.vals, chr.lengths[,c('chrom','chr.length','genome.length')])
  df$g.start = df$start.pos; df$g.end = df$end.pos
  df[which(df$chrom != 1), 'g.start'] = rowSums(df[which(df$chrom != 1), c('start.pos', 'genome.length')])
  df[which(df$chrom != 1), 'g.end'] = rowSums(df[which(df$chrom != 1), c('end.pos', 'genome.length')])

  df = melt(df, measure.vars=grep('^D',colnames(df),value=T))

  p1 = p1 + geom_segment(data=df, aes(x=g.start,xend=g.end,y=value, yend=value,color=variable),show.legend=F)  
}

grid.arrange(p, p1, nrow=2)

## Quick look at our demographic information

```{r echo=F}

ggplot(patient.info, aes(Age.at.diagnosis, fill=Status)) + geom_histogram(binwidth=3) + labs(title='Age at diagnosis')

ggplot(patient.info, aes(BMI, fill=Status)) + geom_histogram(binwidth=1) + labs(title='BMI')

ggplot(patient.info, aes(Maximal,Circumference, color=Status)) + geom_point() + labs(title='C vs M')

grid.arrange(ggplot(patient.info, aes(p53.Status, fill=Status)) + geom_bar(position = position_dodge()) + labs(title='p53 Staining'),
  tableGrob(  rbind( 'Non-progressor'=summary(subset(patient.info, Status == 'NP')$p53.Status),
                     'Progressor'=summary(subset(patient.info, Status == 'P')$p53.Status) ) ))

p53Path = table(subset(patient.info, Status == 'P' & p53.Status == 1)$Pathology)
```

`r pander(p53Path/table(subset(patient.info, Status == 'P' & !is.na(p53.Status))$Pathology), caption='Ratio of progressors predicted by p53 staining per pathology where p53 staining was done (excluding N/As)')`

@@

```{r labelsPNP, echo=F, message=F, include=T}
## binomial: prog 1, np 0
labels = unlist(sapply(patient.data, function(df) {
  df$info = arrange(df$info, Endoscopy.Year, Pathology)
  label = as.integer(df$info$Status == 'P') #as.integer(df$info$Pathology %in% c('HGD', 'IMC'))
  names(label) = df$info$Samplename
  return(label)
}))
names(labels) = sub('.*\\.', '',  names(labels))

@@

dysplasia.df = t(mergedDf[,intersect(names(labels), colnames(mergedDf))])
labels = labels[intersect(names(labels), colnames(mergedDf))]
dim(dysplasia.df)

@@

```{r, echo=F, message=F, warning=F}

source('lib/cv-pt-glm.R')

@@

coefs = list(); plots = list(); performance.at.1se = list(); models = list(); cvs = list()
folds = 10; splits = 5 
alpha.values = c(0, 0.5,0.7,0.8,0.9,1)

file = paste(data, 'Analysis/patient_folds.tsv', sep='/')
if (file.exists(file)) {
  message(paste("Reading folds file", file))
  sets = read.table(file, header=T, sep='\t')
} else {
  sets = create.patient.sets(pts, folds, splits, 0.2)  
  write.table(sets, quote=F, sep='\t', row.names=F, file=file)
  stop("Missing patient folds file")
}

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=T)
} else {
  for (a in alpha.values) {
    fit0 <- glmnet(dysplasia.df, labels, alpha=a, nlambda=nl, family='binomial', standardize=F) # all patients

    l = fit0$lambda
    #if (more.l) l = more.l(fit0$lambda)

    cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=l, pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', opt=-1, standardize=F)

    lambda.opt = cv.patient$lambda.1se

@@

    performance.at.1se[[as.character(a)]] = subset(cv.patient$lambdas, `lambda-1se`)
    models[[as.character(a)]] = fit0
    cvs[[as.character(a)]] = cv.patient
  }
  save(plots, coefs, performance.at.1se, dysplasia.df, models, cv.patient, labels, file=file)
}

@@

if (file.exists(file)) {
  load(file, verbose=T)
  # Fix info
  pg.samp = lapply(pg.samp, function(df) {
    merge(df[,c('Samplename',setdiff(colnames(df), colnames(patient.info))) ], subset(patient.info, Patient == unique(df$Patient)), by='Samplename')
  })

} else {
  performance.at.1se = c(); coefs = list(); plots = list(); fits = list()
  # Remove each patient (LOO)
  for (pt in names(pg.samp)) {
    print(pt)
    tmp.patient.data = patient.data[subset(sum.patient.data, Hospital.Research.ID != pt)$Hospital.Research.ID]
    samples = as.vector(unlist(sapply(tmp.patient.data, function(df) df$info$Samplename )))

@@

    # Fit generated on all samples, including HGD
    a = select.alpha
    fitLOO <- glmnet(training, labels[train.rows], alpha=a, family='binomial', nlambda=nl, standardize=F) # all patients
    l = fitLOO$lambda
    cv = crossvalidate.by.patient(x=training, y=labels[train.rows], lambda=l, 
                                  pts=subset(pts, Samplename %in% samples), a=a, nfolds=folds, splits=splits, fit=fitLOO, standardize=F)
    #plots[[pt]] = cv$plot + ggtitle(pt)

@@

ggplot(as.data.frame(performance.at.1se), aes(y=performance.at.1se, x='LOO')) + ylim(0.5,0.9) +
  geom_boxplot( fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA) + geom_jitter() +
  labs(y='performance at lambda.1se', x='', title=paste('Performance for Leave One (patient) Out, at alpha=',select.alpha,sep=''))

@@

@@

```{r, eval=T}

file = paste(cache.dir, 'chr17.predictive.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=T)
} else {
  performance.at.1se = list(); coefs = list()
  for ( d in c(1:22) ) {
    print(d)
    df = dysplasia.df[,grep(paste('^',d,':',sep=''), colnames(dysplasia.df))]
    fit = run.fit(df, labels)
    performance.at.1se[[ paste('chr', d, sep='') ]] = fit$perf
    coefs[[ paste('chr', d, sep='') ]] = fit$coefs
  }

  ## No HGD
  df17 = dysplasia.df[,grep('^17:', colnames(dysplasia.df))]
  df17nohgd = df17[!is.na(match(rownames(dysplasia.df), subset(patient.info, Pathology %in% c('BE','ID','LGD'))$Samplename)),]
  fit = run.fit(df17nohgd, labels[rownames(df17nohgd)])
  performance.at.1se[['chr17.noHGD']] = fit$perf
  coefs[['chr17.noHGD']] = fit$coefs

  save(performance.at.1se, coefs, file=file)
}
